---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rallicagram

<!-- badges: start -->
<!-- badges: end -->

## Overview

`rallicagram` is a R wrapper for the [Gallicagram API](https://regicid.github.io/api). 

[Gallicagram](https://shiny.ens-paris-saclay.fr/app/gallicagram) is a super nice tool and set of databases that enables to easily run simple Natural Language Processing (**NLP**) analyses on a wide set of corpora. In particular, it enables to build historical **time series of occurrences of keywords** in various media in one line of code.

## Installation

You can install the development version of rallicagram from [GitHub](https://github.com/vincentbagilet/rallicagram) with:

``` r
# install.packages("devtools")
devtools::install_github("vincentbagilet/rallicagram")
```

## Usage

The main function, `gallicagram`, builds a data frame with the yearly, monthly or daily proportion of mentions of a term.

```{r usage}
library(rallicagram)

gallicagram(keyword = "président", corpus = "lemonde", from = 1960)
```

It enables to draw nice graphs representing the evolution of the use of a term in time, in two lines of code.

```{r graph, dpi=300, out.width="70%", fig.align = "center", echo=FALSE, message=FALSE}
library(mediocrethemes)
set_mediocre_all()

gallicagram(keyword = "obama", from = 2005) |> 
  gallicagraph()
```

The package also allows to describe co-occurrences or words associated with a keyword. The corresponding functions are described in the [vignette](articles/rallicagram.html).

## Corpora

The corpora available via Gallicagram are:

```{r list_corpora, echo=FALSE}
rallicagram::list_corpora |> 
  dplyr::rename_with(\(x) stringr::str_replace(x, "_", " ") |>  
                       stringr::str_to_title()) |> 
  knitr::kable()
```

Additional information on Gallicagram can be found on a [preprint](https://osf.io/preprints/socarxiv/84bf3/) by Gallicagram developers [
Benoît de Courson](https://regicid.github.io/) and Benjamin Azoulay and on the "Notice" tab of the [Gallicagram website](https://shiny.ens-paris-saclay.fr/app/gallicagram).
