---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rallicagram

<!-- badges: start -->
<!-- badges: end -->

## Overview

`rallicaggram` calls the Gallicagram API directly from R. [Gallicagram](https://shiny.ens-paris-saclay.fr/app/gallicagram) enables to build time series of keywords used in a set of French corpora:

- **Historical newspapers** from [Gallica](https://gallica.bnf.fr/) (corpus="press"). A corpus of 3 million issues, reliable from 1789 to 1950. Yearly or monthly resolution.
- **Public domain books** available on [Gallica](https://gallica.bnf.fr/) (corpus="books"). 300 000 books, more and more reliable through the 16th to 18th centuries and up to 1950. Yearly resolution.
- **Articles published in [*Le Monde*](https://www.lemonde.fr/)** (corpus="lemonde"). Available and reliable from December 1944 to February 22, 2022. Yearly, monthly or daily resolution.

Additional information on a [preprint](https://osf.io/preprints/socarxiv/84bf3/) by Gallicagram developers [
Benoît de Courson](https://regicid.github.io/) and [Benjamin Azoulay](https://benjamin-azoulay.my.canva.site/) and on the "Notice" tab of the [Gallicagram website](https://shiny.ens-paris-saclay.fr/app/gallicagram).

## Installation

You can install the development version of rallicagram from [GitHub](https://github.com/vincentbagilet/rallicagram) with:

``` r
# install.packages("devtools")
devtools::install_github("vincentbagilet/rallicagram")
```

## Usage

The main function, `gallicagram`, builds a data frame with the yearly, monthly or daily proportion of mentions of a term in one of the three corpora between two specified dates.

```{r usage}
library(rallicagram)

gallicagram(
  keyword = "président", 
  corpus = "lemonde", 
  from = 1960, 
  to = 1970,
  resolution = "monthly"
)
```

It enables to draw nice graphs representing the evolution of the use of a term in time.

```{r graph, dpi=300, out.width="85%", fig.align = "center"}
library(mediocrethemes)
set_mediocre_all()

ex_data <- gallicagram(
  keyword = "informatique", 
  corpus = "lemonde", 
  from = 1945, 
  to = 2022,
  resolution = "yearly"
)

ex_data |> 
  ggplot(aes(x = date, y = prop_occur)) + 
  geom_line(size = 0.7) +
  labs(
    title = "Evolution of occurrences of 'computering' (informatique)",
    subtitle = "In the french newspaper Le Monde",
    x = NULL,
    y = "Proportion of occurrences"
  )
  
```


Additional functions, to describe close co-occurrences for instance, are also available in this package and described in the [vignette](articles/rallicagram.html).
