---
title: "rallicagram"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rallicagram}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, message=FALSE, warning=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "70%",
  dpi = 400,
  fig.align = "center",
  dev = "png",
  fig.width = 7,
  fig.height = 5
)
```

```{r packages}
library(rallicagram)
```

# Main functions

## Occurrences

The main function in this package, `gallicagram`, builds a data frame with the yearly, monthly or daily proportion of mentions of a term in one of the three corpora between two specified dates.

```{r ex_occur}
ex_occur <- gallicagram(
  keyword = "président",
  corpus = "lemonde",
  from = 1960,
  to = 1970,
  resolution = "monthly"
)
```

Here is an example of part of the output:

```{r kable_occur, echo=FALSE}
ex_occur |>
  utils::head() |>
  knitr::kable()
```

## Co-occurrences

`gallicagram_cooccur` builds a data frame with the yearly, monthly or daily proportion of mentions of close co-occurrences of two keywords in one of the three corpora between two specified dates.

Close co-occurrences correspond to the number of 3-grams (4-grams in the *Le Monde* corpus) that contain the two keywords.

The parameter `keywords` takes a character vector of length two as input.

```{r ex_cooccur}
ex_cooccur <- gallicagram_cooccur(
  keywords = c("président", "république"),
  corpus = "lemonde",
  from = 1960,
  to = 1970,
  resolution = "monthly"
)
```

Here is an example of part of the output:

```{r kable_cooccur, echo=FALSE}
ex_cooccur |>
  utils::head() |>
  knitr::kable()
```

## Associations

`gallicagram_associated`  builds a data frame the words the most frequently used following a given ngram over the period. 

For instance "camarade" is often followed by "staline" or "khrouchtchev" in Le Monde. The function returns the most frequent ngrams of the form "camarade \*". Setting `after = FALSE` also includes the most frequent ngrams of the form "\* camarade".

The keyword can be a 2-gram in the "books" and "press" corpora and a 3-gram in the "lemonde" corpus. Searching the "press" corpus can require a long running time.

```{r ex_assoc}
ex_associated <- gallicagram_associated(
  keyword = "camarade",
  corpus = "lemonde",
  from = 1960,
  to = 1970
)
```

Here is an example of part of the output:

```{r kable_assoc, echo=FALSE}
ex_associated |>
  utils::head() |>
  knitr::kable()
```

# Searching several keywords

In order to search several keywords at once, we can use the function `purrr::map_dfr`. It first takes as parameters the vector of keywords to search, then the `gallicagram` function followed by additional arguments to pass to the `gallicagram` function. It returns a unique data frame with all the results, basically binding the two tibbles produced by each search.

```{r map, eval=FALSE}
library(purrr)

keywords <- c("république", "france")

purrr::map_dfr(keywords, gallicagram,
               corpus = "lemonde", from = "1960", to = "1970")
```

Note that based on the same principle, we can run the function for a set of various parameters at once. We only need to specify it in a data frame and can pass it to `purrr::pmap_dfr`. One can define the set of parameters in a tibble and use the function `tidyr::crossing` do create all combination of possible searches.

```{r params_pmap}
params_pmap <-
  tibble::tibble(
    from = 1850,
    to = 1870,
    resolution = "yearly"
  ) |>
  tidyr::crossing(corpus = c("press", "books")) |>
  tidyr::crossing(keyword = c("république", "france"))
```

The set of parameters to search looks like this:

```{r kable_params_pmap}
knitr::kable(params_pmap)
```

We can then pass it to the function:

```{r, eval=FALSE}
purrr::pmap_dfr(params_pmap, gallicagram)
```
