---
title: "rallicagram"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rallicagram}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, message=FALSE, warning=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "70%",
  dpi = 400,
  fig.align = "center",
  dev = "png",
  fig.width = 7,
  fig.height = 5
)
```

```{r packages, echo=FALSE}
library(rallicagram)
```

`rallicagram` is a R wrapper for the [Gallicagram API](https://regicid.github.io/api). More info on Gallicagram and its API can be found on a [preprint](https://osf.io/preprints/socarxiv/84bf3/) by Gallicagram developers [
Benoît de Courson](https://regicid.github.io/) and Benjamin Azoulay, on the "Notice" tab of the [Gallicagram website](https://shiny.ens-paris-saclay.fr/app/gallicagram) and on the [API documentation](https://regicid.github.io/api).

# Main functions

## Occurrences

The main function in this package, `gallicagram`, builds a data frame with the yearly, monthly or daily proportion of mentions of a term in one of the three corpora between two specified dates. 

This function corresponds to the `Query` route of the [API](https://regicid.github.io/api).

```{r ex_occur}
ex_occur <- gallicagram(
  keyword = "président",
  corpus = "lemonde",
  from = 1960,
  to = 1970,
  resolution = "monthly"
)
```

Here is an example of part of the output:

```{r kable_occur, echo=FALSE}
ex_occur |>
  utils::head() |>
  knitr::kable()
```

### Occurences for all copora and the longest period possible

The function `gallicagram_all` enables to retrieve the yearly proportion of occurrences of a keyword in each of the corpora for the longest time range for which the data is available and reliable. 

It only takes one parameter: the keyword of interest. It then calls `gallicagram(keyword)` for each corpus and for the longest period possible for the corpus and returns a data frame with the outcome of each call.

```{r ex_all, eval=FALSE}
gallicagram_all("président")
```

## Co-occurrences

`gallicagram_cooccur` builds a data frame with the yearly, monthly or daily proportion of mentions of close co-occurrences of two keywords in one of the three corpora between two specified dates.

Close co-occurrences correspond to the number of 3-grams (4-grams in the *Le Monde* corpus) that contain the two keywords. This function corresponds to the `Contain` route of the [API](https://regicid.github.io/api).

The parameters `keyword_1` and `keyword_2` take the words to look co-occurrences for.

```{r ex_cooccur}
ex_cooccur <- gallicagram_cooccur(
  keyword_1 = "président",
  keyword_2 = "république",
  corpus = "lemonde",
  from = 1960,
  to = 1970,
  resolution = "monthly"
)
```

Here is an example of part of the output:

```{r kable_cooccur, echo=FALSE}
ex_cooccur |>
  utils::head() |>
  knitr::kable()
```

## Associations

`gallicagram_associated`  builds a data frame with the words most frequently used close (at a specified distance) to a given keyword over the period. This function calls the `associated` route of the [API](https://regicid.github.io/api).

The keyword can be a 2-gram in the "books" and "press" corpora and a 3-gram in the "lemonde" corpus. The length of the keyword plus the distance cannot be more than 3 for the "books" and "press" corpora and a 4 for the "lemonde" corpus. Searching the "press" corpus can take a long time. 

```{r ex_assoc}
ex_associated <- gallicagram_associated(
  keyword = "camarade",
  corpus = "lemonde",
  from = 1960,
  to = 1970,
  n_results = 10,
  distance = 3
)
```

The terms most often associated with "camarade" are:

```{r kable_assoc, echo=FALSE}
ex_associated |>
  knitr::kable()
```

By default, the most frequent words in the Gallica books corpus are removed. Instead, we can remove other stopwords, by supplying a list of words to ignore to the `stopwords` parameter of the function. For instance, we can use the list in the `lsa` package, `lsa::stopwords_fr`. To keep the stopwwords, set the `stopwords` parameter to `NULL`.

```{r ex_assoc_stopwords}
ex_associated_stopwords <- gallicagram_associated(
  keyword = "camarade",
  corpus = "lemonde",
  from = 1960,
  to = 1970,
  n_results = 10,
  stopwords = NULL
)
```

The most often associated terms without excluding stopwords are:

```{r kable_assoc_stopwords, echo=FALSE}
ex_associated_stopwords |>
  utils::head() |>
  knitr::kable()
```

# More than one keyword

## Searching several keywords

In order to run the functions for several keywords at once, we can use the function `purrr::map_dfr`. It takes as parameters the vector of keywords to search, followed by the function (`gallicagram`) and additional arguments to pass to the function. It returns a unique data frame with the results for searches corresponding to each keyword, basically binding the dataframes produced by each keyword search.

```{r map, eval=FALSE}
library(purrr)

keywords <- c("république", "france")

purrr::map_dfr(keywords, gallicagram,
               corpus = "lemonde", from = "1960", to = "1970")
```

Based on that same principle, we can run several searches at once, varying all parameters, not only the keyword searched. To do that, we can specify the series of parameters in a data frame, each row corresponding to a set of parameters to run the `gallicagram` function for. We then pass this data frame to `purrr::pmap_dfr`. 

To specify the set of parameters, we can either build the parameters data frame by hand. It is also often helpful to use `tidyr::crossing` to create all combination of possible searches.

```{r params_pmap}
params_pmap <-
  tibble::tibble(
    from = 1850,
    to = 1870,
    resolution = "yearly"
  ) |>
  tidyr::crossing(corpus = c("press", "books")) |>
  tidyr::crossing(keyword = c("république", "france"))
```

The corresponding set of parameters to search looks like this:

```{r kable_params_pmap, echo=FALSE}
knitr::kable(params_pmap)
```

We can then pass it to `purrr::pmap_dfr` that will call the function `gallicagram` for each of the 4 sets of parameters defined in the rows of `params_pmap`:

```{r pmap, eval=FALSE}
purrr::pmap_dfr(params_pmap, gallicagram)
```

This method also apply for the other functions in the `rallicagram` package.

## Occurrences of a lexicon

If one wants to compute occurrences in a lexicon, co-occurrences in two lexicons or words associated with a lexicon, one can use the functions with suffix `_lexicon` (`gallicagram_lexicon`, `gallicagram_cooccur_lexicon` and `gallicagram_associated_lexicon`). They simply loop the function over each  word the lexicon(s) and sums the results. It takes as inputs vectors of keywords:

```{r ex_lexicon}
ex_lexicon <- gallicagram_lexicon(
  lexicon = c("président", "présidente"),
  corpus = "lemonde",
  from = 1960,
  to = 1970,
  resolution = "monthly"
)

ex_cooccur_lexicon <- gallicagram_cooccur_lexicon(
  lexicon_1 = c("président", "présidente"),
  lexicon_2 = c("mauvais", "nul"),
  corpus = "lemonde",
  from = 1960,
  to = 1970,
  resolution = "monthly"
)
```

Here is an example of part of the output of `gallicagram_lexicon`:

```{r kable_lexicon, echo=FALSE}
ex_lexicon |>
  utils::head() |>
  knitr::kable()
```

## Words with the same stem

The function `get_same_stem()` takes as input a keyword and returns a vector of words that share the same stem. For instance, for "écologie":

```{r get_same_stem}
get_same_stem("écologie")
```

This function is particularly useful when its output is passed to one of the `_lexicon` functions:

```{r ex_stem_lexicon}
ex_stem_lexicon <- get_same_stem("écologie") |> 
  gallicagram_lexicon(
    corpus = "lemonde",
    from = 1960,
    to = 1970,
    resolution = "yearly"
  )
```

It enables to get all the occurrences for words that share the same stem as the keyword. Note that due to the number of requests, it may take some time to run. 

```{r kable_stem_lexicon, echo=FALSE}
ex_stem_lexicon |>
  utils::head() |>
  knitr::kable()
```

# Graphs

One of the main usage of `Gallicagram` is to plot time series of occurrences in a corpus. Here is an example graph using `ggplot` (and my own ggplot theme [`mediocrethemes`](https://vincentbagilet.github.io/mediocrethemes/). The use of this theme is of course not mandatory and one can remove the line `mediocrethemes::set_mediocre_all()`): 

```{r raw_ts, message=FALSE}
library(ggplot2)
mediocrethemes::set_mediocre_all()

ex_data <- gallicagram(
  keyword = "informatique",
  corpus = "lemonde",
  from = 1945,
  to = 2022,
  resolution = "yearly"
)

ex_data |>
  ggplot(aes(x = date, y = prop_occur)) +
  geom_line() +
  labs(
    title = "Evolution of occurrences of 'computering' (informatique)",
    subtitle = "In the french newspaper Le Monde",
    x = NULL,
    y = "Proportion of occurrences"
  )
```

We can use `geom_smooth` to plot smoother lines. When using the default loess smoothing, the parameter `span` controls the amount of smoothing. Larger values are associated with less smoothing. Here is the same graph but smoothed:

```{r smoothed_ts}
ex_data |>
  ggplot(aes(x = date, y = prop_occur)) +
  geom_smooth(span = 0.2) +
  labs(
    title = "Evolution of occurrences of 'computering' (informatique)",
    subtitle = "In the french newspaper Le Monde",
    x = NULL,
    y = "Proportion of occurrences"
  )
```
