---
title: "rallicagram"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rallicagram}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, message=FALSE, warning=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "70%",
  dpi = 400,
  fig.align = "center",
  dev = "png",
  fig.width = 7,
  fig.height = 5
)
```

```{r packages, echo=FALSE}
library(rallicagram)
```

This vignette introduces the main functions in the package and provides examples of simple analyses that can be carried out with these functions. 

More information on each function is available in the documentation of each function, accessible by typing in the console `help("gallicagram")` for the `gallicagram` function instance.

# Main functions

## Occurrences

The main function in this package, `gallicagram`, builds a data frame with the yearly, monthly or daily proportion of mentions of a term in one of the corpora. 

For the *Le Monde* corpus, this function allows to compute the number of newspaper articles that mention the keywords (parameter `n_of = "articles"`). 

```{r ex_occur}
ex_occur <- gallicagram(
  keyword = "président",
  corpus = "lemonde",
  from = 1960,
  to = 1970,
  resolution = "monthly"
)
```

Here is an example of the first rows of the output:

```{r kable_occur, echo=FALSE}
ex_occur |>
  utils::head() |>
  knitr::kable()
```

## Co-occurrences

`gallicagram_cooccur` builds a data frame with the yearly, monthly or daily proportion of mentions of close co-occurrences of two keywords in one of the three main corpora (`lemonde`, `livres` and `presse`) between two specified dates.

Close co-occurrences correspond to words that are less than 3 (4 in the *Le Monde* corpus) words away in the initial text. For the *Le Monde* corpus, this function allows to compute the number of co-occurrences within an entire article (parameter `cooccur_level = "articles"`). 

```{r ex_cooccur}
ex_cooccur <- gallicagram_cooccur(
  keyword_1 = "président",
  keyword_2 = "république",
  corpus = "lemonde",
  from = 1960,
  to = 1970,
  resolution = "monthly"
)
```

Here is an example of part of the output:

```{r kable_cooccur, echo=FALSE}
ex_cooccur |>
  utils::head() |>
  knitr::kable()
```

## Associations

`gallicagram_associated` builds a data frame with the words most frequently used close  to a given keyword over the period. For the *Le Monde* corpus, this function allows to compute the number of co-occurences within an entire article (parameter `distance = "articles"`). 

```{r ex_assoc}
ex_associated <- gallicagram_associated(
  keyword = "camarade",
  corpus = "lemonde",
  from = 1960,
  to = 1970,
  n_results = 10,
  distance = 3
)
```

The terms most often associated with "camarade" are:

```{r kable_assoc, echo=FALSE}
ex_associated |>
  knitr::kable()
```

# Graphs

One of the main usage of `Gallicagram` is to plot time series of occurrences in a corpus. The function `gallicagraph` enables to do that in only one additional line of code. It take as input a dataframe produced by one of the functions from the package and automatically produces the corresponding graph:

```{r graph_line, message=FALSE}
# my own theme for ggplot graphs
mediocrethemes::set_mediocre_all()

gallicagram(keyword = "obama", corpus = "lemonde", from = 2005) |> 
  gallicagraph()
```

Note that I use my own ggplot theme [`mediocrethemes`](https://vincentbagilet.github.io/mediocrethemes/). The use of this theme is of course not mandatory and one can remove the line `mediocrethemes::set_mediocre_all()`).

One can also use the same function to plot words associated with a given keyword:

```{r graph_associated, message=FALSE, fig.asp=1.2}
gallicagram_associated(keyword = "obama", corpus = "lemonde", from = 2005) |> 
  gallicagraph()
```

# More than one keyword

## Occurrences of a lexicon

Each of the main function described above has a `lexicon` counterpart, adding a suffix `_lexicon` (`gallicagram_lexicon`, `gallicagram_cooccur_lexicon` and `gallicagram_associated_lexicon`). Each function simply loop the intial function over each word the lexicon(s) and sums the results. It thus compute the sum of (co-)occurrences of each word in the lexicon:

```{r ex_lexicon}
ex_lexicon <- gallicagram_lexicon(
  lexicon = c("président", "présidente"),
  corpus = "lemonde",
  from = 1960,
  to = 1970,
  resolution = "monthly"
)

ex_cooccur_lexicon <- gallicagram_cooccur_lexicon(
  lexicon_1 = c("président", "présidente"),
  lexicon_2 = c("mauvais", "nul"),
  corpus = "lemonde",
  from = 1960,
  to = 1970,
  resolution = "monthly"
)
```

Here is an example of part of the output of `gallicagram_lexicon`:

```{r kable_lexicon, echo=FALSE}
ex_lexicon |>
  utils::head() |>
  knitr::kable()
```

These functions are particularly helpful to compute the number of occurrences of words from a whole lexical field (or the plural or feminine form of a keyword, etc).

## Words with the same stem

The function `get_same_stem()` enables to retrieve part of these forms: those that share the same stem as the keyword of interest. For instance, for "écologie":

```{r get_same_stem}
get_same_stem("écologie")
```

This function is particularly useful when its output is passed to one of the `_lexicon` functions:

```{r ex_stem_lexicon}
ex_stem_lexicon <- get_same_stem("écologie") |> 
  gallicagram_lexicon(
    corpus = "lemonde",
    from = 1960,
    to = 1970,
    resolution = "yearly"
  )
```

Note that due to the number of requests sent to the API, it may take some time to run. 

```{r kable_stem_lexicon, echo=FALSE}
ex_stem_lexicon |>
  utils::head() |>
  knitr::kable()
```

## Searching several keywords

One can also run the functions for several keywords but separately, without suming the result. To do so, one can use the function `purrr::map_dfr`. It takes as parameters the vector of keywords to search, followed by the function (`gallicagram`) and additional arguments to pass to the function. It returns a unique data frame with the results for searches corresponding to each keyword, basically binding the dataframes produced by each keyword search.

```{r map, eval=FALSE}
library(purrr)

keywords <- c("république", "france")

purrr::map_dfr(keywords, gallicagram,
               corpus = "lemonde", from = "1960", to = "1970")
```

Based on that same principle, we can run several searches at once, varying all parameters, not only the keyword searched. To do that, we can specify the series of parameters in a data frame, each row corresponding to a set of parameters to run the `gallicagram` function for. We then pass this data frame to `purrr::pmap_dfr`. 

To specify the set of parameters, we can either build the parameters data frame by hand. It is also often helpful to use `tidyr::crossing` to create all combination of possible searches.

```{r params_pmap}
params_pmap <-
  tibble::tibble(
    from = 1850,
    to = 1870,
    resolution = "yearly"
  ) |>
  tidyr::crossing(corpus = c("press", "books")) |>
  tidyr::crossing(keyword = c("république", "france"))
```

The corresponding set of parameters to search looks like this:

```{r kable_params_pmap, echo=FALSE}
knitr::kable(params_pmap)
```

We can then pass it to `purrr::pmap_dfr` that will call the function `gallicagram` for each of the 4 sets of parameters defined in the rows of `params_pmap`:

```{r pmap, eval=FALSE}
purrr::pmap_dfr(params_pmap, gallicagram)
```

This method also applies for the other functions in the `rallicagram` package.
